Self Referencing Sequences

### 2014-10-31 draft.tex
sdfh

In this paper, we address a variety of questions regarding these sequences. In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. After that, we consider how the starting number for the sequence affects the long term behavior of the sequence. In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers. In the following Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

@@

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number. To see why this is true, we give an algorithm to build the sequence. Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$. If we ever write down some $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
\begin{quote}
\textbf{Algorithm:}\\
Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that block. We read that second element, and then we add a block of $a_i$ instances of $a_{i+1}$ to the end of that sequence. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\

This procedure continues deterministically so long as there is an element to read. Every time we write down a block, we read 1 element and write down at least 1 element, thus the end of the sequence will always be at or after the spot we are reading.
\end{quote}

Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number. From this point on, we will denote by $s(G,g)$ the unique sequence generated by set $G$ with starting number $g \in G$.

\section{Prefixes}
\label{prefixes}
\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
\begin{definition} We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a prefix of length $n$ if 
\begin{equation*}
a_{n+i} = b_{i},
\end{equation*}
for $i$ any positive integer. We say that two sequences are independent if they do not differ by a prefix for any $n$, in either order. 
\end{definition}
Note that a pair of sequences can differ by multiple prefixes of different length. For example if we have 
\begin{align*}
&\{a_i\} = \{1,2,3,4,5,6,1,2,3,4,5,6, \ldots \} \\  
&\{b_i\}= \{7,8,9,1,2,3,4,5,6,1,2,3,4,5,6, \ldots\},
\end{align*}
 then $a_i$ and $b_i$ differ by prefixes of length $3+6k$ for $k$ a nonnegative integer. Because of this we make the following definition.
 \begin{definition} 
 We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a minimal prefix of length $n$ if they differ by a prefix of length $n$, but do not differ by a prefix of length $m$ for all $0<m<n$.
 \end{definition}
We note that if a pair of sequences differs by a prefix they differ by a minimal prefix. In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$. The following lemma is a direct consequence of our algorithm for generating self-referential sequences:

\begin{lemma} Let $G = \{1, g_1, g_2, \ldots, g_n\}$, $s(G, 1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
\end{lemma}
\begin{proof}
Let $a_i$ refer to the $i$th term of $s(G, 1)$ and $b_i$ refer to the $i$th term of $s(G, g_1)$. If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = g_1$, and that this is the beginning of a new block of length $g_1$. Furthermore, $b_1 = g_1$ and it indicated the beginning of a block of length $g_1$. Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements. Because the number and the begining of the block is the same. The algorithm will generate the same sequence, shifted by 1. Therefore, $s(G,1)$ and $s(G, g_1)$ differ by a minimal prefix of length at most 1.
\end{proof}
We now give a 
\begin{theorem}
\label{independentsequences}
Let $G = \{ 1, a, b\}$ where $1<a<b$. $s(G, a)$ and $s(G, b)$ are independent.
\end{theorem}
\begin{proof}
We proceed by contradiction. If the two sequences are not independent then they must differ by some minimal prefix of length $n$. If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ then
\begin{equation*}
a_{n+k} = b_{k}.
\end{equation*}
We know that the first $b$ terms of $b_i$ are $b$'s. Because $b>a>1$ no block can be longer then $b$. Thus the terms immediately preceding and following this block in  $a_i$ cannot be $b$'s. Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, for some $m$ a positive integer. Since the value of the sequence at position $k$ is the length of the $k$th block this means that $a_{m+1}=b_1$. In fact the $k$th block of $b_i$ is the $(m+k)$th block of $a_i$ and so 
\begin{equation*}
a_{m+k} = b_k
\end{equation*}

We now claim that the prefix has length strictly larger than $m$. Assume otherwise so $n \leq m$. 
This means that $a_{m+1}$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block, which begins at position $a_{n+1}$. By our proof that self-referential sequences are well defined we know that the second case cannot occur. Therefore we must have the $m+1$th block start at position $m+1$. This would require every block to have length exactly 1, but this is a contradiction because the first block has length $a_1=a>1$. Therefore the prefix must have length strictly larger than $m$.

@@

In this section, we consider a slightly weaker definition of equivalence. Consider, for example, sequences generated by $G = \{1,2,3\}$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases. Lets look at the state of the algorithm that generates both sequences. After the first block, the execution of the algorithm of $s(G, 2)$ right after placing the first $3$'s (in the 3rd and 4th position), the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements in the current block
\item starting number is $1$
\end{itemize}.   
If we look at the execution of the algorithm that generates $s(G, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements in the current block
\item starting number is $1$
\end{itemize}
Therefore, since the states are the same, the sequence generated after the 4th position of $s(G, 2)$ and after the $3rd$ position of $s(G, 3)$ will be the same. By inspection, we see we can also include two of the 3's.

Thus, since both sequences are essentially the same if a prefix is removed from each of them, we call these sequences \emph{equivalent}. To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}

@@

This conjecture is partly motivated by the fact that the property holds for $\{1,2,3\}$. For generating sets with larger numbers, we have been unable to verify or refute the conjecture. However, in Section~\ref{density}, we show that for some generating sets the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $a$. We define the read distance for an element $a_i$ as the number of terms in between it and the beginning of the $i$th block of numbers.
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $a_i$ generated by $\{1,2,3\}$ beginning with a 2  
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$. From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing. We can in fact compute the read distance exactly. 
\begin{lemma} 
\label{readdistformula}
Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$, for $i>1$, is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. When $i=2$ we must show that the read distance is $a_1-1$. The first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at position $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the start of the $l$th block and position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $(l+1)$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and so the read distance is monotonically increasing. Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a simple bound on the number of consecutive times the read distance can stay constant.

\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max(a_1, a_2, \ldots, a_n)$. Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$. 
\end{lemma}
\begin{proof}
Assume otherwise. By Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$. Thus it must be exactly $d$. Lemma ~\ref{readdistformula} also tells us that if this is the case then $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$. But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence. Therefore the read distance for $x_{i}$ is at least $d+1$. 
\end{proof}

@@

One possible iterative process is as follows. Let the sequence $s^0$ be 2, and let the sequence $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$. This gives:
\begin{align*}
s^{(0)}&= 2\\
s^{(1)} &= 22\\
s^{(2)} &= 2211\\
s^{(3)}&= 221121\\
s^{(4)} &= 221121221\\
&\vdots
\end{align*}
Let $s^\infty$ be the limit of this process. 
By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as $i \rightarrow \infty$. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2. In fact we can prove a more general result about this iterative process.

\begin{theorem}
Let $s^0 = a$ for some $a \in A$ where $a \neq 1$, and let $s^{(i)}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $s^{(i-1)}$. Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
\end{theorem}

\begin{proof}
We prove this with induction. We show that if $s^{(i)}$ is a prefix of $s(A,a)$, then $s^{(i+1)}$ is a longer prefix of $s(A,a)$. The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(A,a)$.

For the inductive step, we observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence. We know that the sequence it describes is longer because the sequence it describes has length equal to the sum of the original prefix. In other words, the length of $s^{(i+1)}$ equals the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1).

The resultant sequence is a prefix of the self referencing sequence because if we apply our process to any prefix of the self referencing sequence it will produce a prefix of the self referencing sequence. 
\end{proof}

Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$.

It turns out that we can be less restrictive about what our starting sequence is. Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2. For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating $s^{(i)}$ give:
\begin{align*}
s^{(0)}&= \mathbf{2},1,1,1,2\\
s^{(1)} &= \mathbf{2,2,1},2,1,2,2\\
s^{(2)} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
s^{(3)} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
&\vdots
\end{align*}
Here, we bold the numbers that form a prefix of $s(\{1,2\},2)$. We note that $s^\infty = s(A,a)$ because because $s^{(0)}$ contains a 2 at the beginning, which by the above theorem will produce $s(A,a)$ in the limit. We call the rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ the \emph{tail}.

We now consider two different forms of convergence. We know that $s^{(i)}$ in some sense approaches $s(A,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(A,a)$ as $i$ increases. However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$. Thus, we formalize two types of convergence
\begin{definition} We say that a sequence $s^{i}$ is weakly convergent to $s(A,a)$ if $s^\infty = s(A,a)$. We say that the sequence $s^{i}$ is strongly convergent if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.
\end{definition}
It is unclear what conditions must be placed on the starting sequence in order to ensure strong convergence, although based on numerical simulation we make the following conjecture.

\begin{conjecture} A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 2)$ if the starting sequence begins with a $2$ and contains no more than 3 elements. A sequence $s^i$ is strongly convergent to $s(\{1,2\}, 1)$ if the starting sequence begins with a $1$ and contains between 2 and 3 elements. A sequence $s^i$ is weakly convergent but not strongly convergent to $s(\{1,2\}, a)$ if the starting sequence begins with an $a \in \{1,2,\}$ and contains 4 or more elements.
\end{conjecture}
We make this conjecture on the basis on computer simulations which executes the iterative process to at least 20 iterations. More iterations were not completed because after this point the simulations began to take impractically long amounts of time to complete.

@@

We would like to know if we can do this for any generating set $G$ and starting number $g$. In particular we would like to be able to simulate our original algorithm with a substitution rule. Previously we established that our algorithm had a state corresponding to 
\begin{itemize}
\item current block size
\item number of elements already written down in the current block
\item what number the current block is composed of 
\end{itemize}

When we apply our substitution rules we should not need to know what happened to blocks other than the one we are currently substituting for. Therefore, each rule must somehow encode the length of  blocks, the number of elements already written in the current block, and what number the current block is composed of. The first condition is simple to implement, as the rules will specify the length of blocks. In order to keep track of the numbers already written, each rule should be self-contained. That is each rule should write complete blocks. In order to comply with the third condition it is enough for the size of the generating set to divide the length of the rules. This is so that at the start of each rule, we always know which numbers to write.

Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating set $G$. We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|G|^{l}$ strings are in the rule set and for each $r \in R$
\begin{itemize}
\item each rule writes blocks in the same order of $G$
\item $\sum_{i \in r} i$ is divisible by $l$
\item $l$ is divisible by $|G|$
\end{itemize}
Given these conditions on every rule, we can guarantee that the self-referential sequence is generated. Note that in this case, the above example does not satisfy our rules until we say that $1, 3 \to 3, 1, 1, 1$.

Another example, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
\begin{align*}
\text{Rule }A: 2,2 &\to 2, 2, 4, 4 \\
\text{Rule }B: 2,4 &\to 2, 2, 4,4,4,4\\
\text{Rule }C: 4,2 &\to 2,2,2,2,4,4\\
\text{Rule }D: 4,4 &\to 2, 2, 2, 2, 4, 4, 4, 4
\end{align*}

@@

We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number. However, generalizing the procedure to improve the bound seems dificult.   
\end{document}

>>>>

In this paper, we address a variety of questions regarding these sequences. In Section~\ref{determinism} we show that if we are given a generating set and a starting number for the sequence, the sequence is uniquely determined. After that, we consider how the starting number for the sequence affects what the sequence can be. In particular, in Section~\ref{prefixes}, we show that there exist limits on how much two sequences with the same generating set can resemble each other if they start with different numbers. In the following Section~\ref{equivalence}, we consider the limiting behavior of such sequences, and we conjecture that the limiting behavior of a sequence is determined only by the generating set, and not the starting number.

@@

We claim that a self-referencing sequence is uniquely determined by the generating set and the starting number. To see why this is true, we can simply give an algorithm to build the sequence. Assume the generating set is $\{a_1,a_2,\dots,a_m\}$ and that we start with some $a_i$. If we ever write down some $a_k$ where $k > m$, we set this equal to $a_{k (\text{mod } m)}$.
\begin{quote}
\textbf{Algorithm:}\\
Suppose the first number of the sequence is $a_i$. Then the first block of the sequence must be a block of $a_i$ instances of $a_i$, so we extend the sequence to this block. If $a_i$ is not 1, then there is a second element of that sequence. We read that second element, and then we add a block to the end of that sequence with length dictated by the value of the second element. If $a_i$ is equal to 1, we write down $a_{i+1}$ in the second spot, and use the value of $a_{i+1}$ to dictate the length of the second block.\\

This procedure can continue deterministically so long as there is always an element to read. However, every time we write down a block, we read 1 element and write down at least 1 element, so the end of the sequence will always be after the spot we are reading. The only exception is when $a_i = 1$, but then since $a_{i+1} \neq 1$, this issue will never occur again.
\end{quote}

Thus, a self-referencing sequence is uniquely determined by the generating set and the starting number. From this point on, we will denote by $s(A,a)$ the unique sequence generated by set $A$ with starting number $a \in A$.

\section{Prefixes}
\label{prefixes}
\emph{Written by Perry Kleinhenz, edited by Fermi Ma and Erik Waingarten}
\begin{definition} We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a prefix of length $n$ if 
\begin{equation*}
a_{n+i} = b_{i},
\end{equation*}
for $i$ any positive integer. if no such $n$ exists then we say that the two sequences are independent.
\end{definition}
Note that a pair of sequences can have prefixes of different length. For example if we have 
\begin{align*}
&\{a_i\} = \{1,2,3,4,5,6,1,2,3,4,5,6, \ldots \} \\  
&\{b_i\}= \{7,8,9,1,2,3,4,5,6,1,2,3,4,5,6, \ldots\},
\end{align*}
 then $a_i$ and $b_i$ differ by prefixes of length $3+6k$ for $k$ a nonnegative integer. Because of this we make the following definition.
 \begin{definition} 
 We say that two sequences $\{a_i\}$ and $\{b_i\}$ differ by a minimal prefix of length $n$ if they differ by a prefix of length $n$, but do not differ by a prefix of length $m$ for all $0<m<n$.
 \end{definition}
We note that if a pair of sequences differs by a prefix they differ by a minimal prefix. In our above example $a_i$ and $b_i$ differ by a minimal prefix of length $3$. The following lemma is a trivial consequence of our algorithm for generating self-referential sequences:

\begin{lemma} Let $C = \{1, c_1, c_2, \ldots, c_n\}$, $s(C, 1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
\end{lemma}
\begin{proof}
Let $a_i$ refer to the $i$th term of $s(C, 1)$ and $b_i$ refer to the $i$th term of $s(C, c_1)$. If we use our algorithm for generating the sequence, we note that $a_1 = 1$, and so $a_2 = c_1$, and that this is the beginning of a new block of length $c_1$. Furthermore, $b_1 = c_1$ and it indicated the beginning of a block of length $c_1$. Note that our algorithm is ``memory-less'' in the sense that it does not maintain any state about previous elements. Because the number and the begining of the block is the same. The algorithm will generate the same sequence, shifted by 1. Therefore, $s(C,1)$ and $s(C, c_1)$ differ by a minimal prefix of length at most 1.
\end{proof}

\begin{theorem}
\label{independentsequences}
Let $C = \{ 1, a, b\}$ where $1<a<b$. $s(C, a)$ and $s(C, b)$ are independent.
\end{theorem}
\begin{proof}
We proceed by contradiction. If the two sequences are not independent then they must differ by some minimal prefix of length $n$. If we let $a_i$ be the sequence beginning with $a$ and $b_i$ be the sequence beginning with $b$ the
\begin{equation*}
a_{n+k} = b_{k}.
\end{equation*}
We know that the first $b$ terms of the sequence which begins with a $b$ are $b$'s. Because $b>a>1$ no block can be longer then $b$. Thus the terms immediately preceding and following this block in the sequence $a_i$ must not be $b$'s. Because of this the first block of $b_i$ is the $m+1$th block of $a_i$, where $m$ is some positive integer. Since the value of the sequence at position $k$ is also the length of the $k$th block this means that $a_{m+1}=b_1$. In fact the $k$th block of $b_i$ is the $(m+k)$th block of $a_i$ and so 
\begin{equation*}
a_{m+k} = b_k
\end{equation*}

We now claim that the prefix has length strictly larger than $m$. Assume otherwise so $n \leq m$. 
This means the entry in the $a_i$ which gives the length of the $m+1$th block occurs at or after the beginning of the $m+1$th block. By our proof that self-referential sequences are well defined we know that the second case cannot occur. Therefore the only other option would be for the $m+1$th block to start at position $m+1$, but this would require every block to have length exactly 1, but this is a contradiction because $a_1=a>1$, that is the first block has length greater than one. Therefore the prefix must have length strictly larger than $m$.

@@

In this section, we consider a slightly weaker definition of equivalence. Consider, for example, sequences generated by $C = \{1,2,3\}$. Theorem~\ref{independentsequences} shows that the sequence starting with 2
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
and the sequence starting with 3
\begin{align*}
3,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
are independent in the sense that neither sequence is exactly contained in the other. However, if we delete the first two numbers of the sequence starting with 2, and the first number of the sequence starting with 3, we get 
\begin{align*}
3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
in both cases. Lets look at the state of the algorithm that generates both sequences. After the first block, the execution of the algorithm of $s(C, 2)$ right after placing the $3$'s (in the 4th and 5th position), the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements on the current block
\item starting number is $1$
\end{itemize}.   
If we look at the execution of the algorithm that generates $s(C, 3)$ after placing the first three $3$'s, the algorithm is in state corresponding to:
\begin{itemize}
\item block size $3$
\item no elements on the current block
\item starting number is $1$
\end{itemize}
Therefore, since the states are the same, the same sequence will be generated after the 4th position of $s(C, 2)$ and after the $3rd$ position of $s(C, 3)$ will be the same. By inspection, we can see we can include $2$ of the 3's.

Thus, since both sequences are essentially the same if certain prefixes are removed, we call these sequences \emph{equivalent}. To formalize, this, we say that two sequences $\{a_i\}$ and $\{b_i\}$ are equivalent if there exists some integers $n,k \geq 0$ such that
\begin{align*}
a_{n+i} = b_{k+i}
\end{align*}

@@

This conjecture is partly motivated by the fact that the property holds for $\{1,2,3\}$. For generating sets with larger numbers, we have been unable to verify or refute the conjecture. However, in Section~\ref{Density}, we show that the starting number of a sequence does not affect the densities of each number, which provides some evidence for this conjecture.

We now introduce a notion which will help us categorize sequences as equivalent. 
\begin{definition} Consider a sequence $x$. We define the read distance for an element $x_i$ as the number of terms in between it and the beginning of the $i$th block of numbers. We denote this as $Rd(x_i)$. 
\end{definition}
The read distance can be any nonnegative integer. For example if we consider the sequence $a_i$ generated by $\{1,2,3\}$ beginning with a 2  
\begin{align*}
2,2,3,3,1,1,1,2,2,2,3,1,2,3,3,1,1,2,2,3,3,3,\dots
\end{align*}
the read distance for the $a_1$ (the first two) is $0$ as the first block of numbers starts at the first number, the read distance for $a_2$ (the second two) is $1$ as the second block of numbers starts at the third number. We can continue to calculate the read distances for this sequence and see that they form a monotonically increasing sequence $0,1,2,4,6,6,6,6,7,8,9,\ldots$. From this we might suspect that the read distance for any self referencing sequence will also be monotonically increasing. We can in fact show a much stronger result. 
\begin{lemma} 
\label{readdistformula}
Let $a_i$ be any self referencing sequence generated by some set of positive integers. Then the read distance for the element $a_i$ for $i>1$ is 
\begin{equation*}
\sum_{k=1}^{i-1}( a_k - 1)
\end{equation*} 
\end{lemma}
\begin{proof}
We will prove this using induction. We know that for $i=1$ the read distance is $0$, as the first block always starts at the first number. For the base case $i=2$ we must show that the read distance is $a_1-1$. Well we know that the first block will be of length $a_1$ since the sequence is self referencing, therefore the second block will begin at \emph{position} $a_1$. The number of elements between position $a_1$ and $2$ is exactly $a_1 -1$.

Now if we assume that our result holds for $i=l$ we would like to show that it holds for $i=l+1$. We know that the distance between the $l$th block and the number at position $l$ is 
\begin{equation*}
\sum_{k=1}^{l-1}( a_k - 1).
\end{equation*} 
This $l$th block has length $a_l$ and so the $l+1$th block will begin $a_l$ positions to the right of the beginning of the $l$th block. Also note that $a_{l+1}$ is one position to the right of $a_l$. Therefore the beginning of the $l+1$th block is 
\begin{equation*}
a_l-1 + \sum_{k=1}^{l-1} (a_k-1) = \sum_{k=1}^{l} (a_k-1)
\end{equation*} 
positions to the right of $a_{l+1}$. But this is exactly the reading distance for $a_{l+1}$ and so we have our result.
\end{proof}
Note that because $a_i$ is generated by positive integers we always have $a_k-1\geq 0$ and the read distance is monotonically increasing. Although this shows us that read distance never decreases it does not tell us how quickly it increases, our next result gives a bound on the number of consecutive times the read distance can stay constant.

\begin{lemma} Consider a self referencing sequence $x$ generated by $\{a_1, a_2, \ldots, a_n\}$ with $a = \max{ a_1, a_2, \ldots, a_n}$. Suppose the read distance for $x_i$ is $d$, then the read distance for $x_{i+a}$ is at least $d+1$. 
\end{lemma}
\begin{proof}
Assume otherwise. By our Lemma ~\ref{readdistformula} we know that the read distance for $x_{i+a}$ is at least $d$. By our assumption it must be exactly $d$. Lemma ~\ref{readdistformula} also tells us that $x_{i+k}=1$ for $k=0,1,2,\ldots, a_n$. But this means that we have a block of length $a+1$, which is a contradiction of the self referencing property of the sequence. Therefore the read distance for $x_{i}$ is at least $d+1$. 
\end{proof}

@@

One possible iterative process is as follows. Let the sequence $\{s^0\}$ be 2, and let the sequence $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with 2 and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$. This gives:
\begin{align*}
\{s^{(0)}\} &= 2\\
\{s^{(1)}\} &= 2,2\\
\{s^{(2)}\} &= 2,2,1,1\\
\{s^{(3)}\} &= 2,2,1,1,2,1\\
\{s^{(4)}\} &= 2,2,1,1,2,1,2,2,1\\
&\cdots
\end{align*}
Let $s^\infty$ be the limit of this process. By limit, we mean the limit with respect to finitely many terms $s^{(i)}_j \rightarrow s^\infty_j$ as$i \rightarrow \infty$. We claim that $s^\infty$ is the unique self-referential sequence $s(\{1,2\},2)$ generated by $\{1,2\}$ with starting number 2.
\begin{proof}
We prove this with induction. We show that if $s^{(i)}$ is a prefix of $s(\{1,2\},2)$, then $s^{(i+1)}$ is a longer prefix of $s(\{1,2\},2)$. The base case of the induction is satisfied, as $s^{(0)}$ is a prefix of $s(\{1,2\},2)$.

For the inductive step, observe that a prefix of the sequence describes the block lengths of a \emph{longer} prefix of the sequence. We know that it describes a longer prefix because the prefix it describes has length equal to the sum of the original prefix. In other words, the length of $s^{(i+1)}$ must be equal to the sum of of the numbers in $s^{(i)}$, which is strictly greater than the length of $s^{(i)}$ (since there are numbers that are greater than 1). This holds true because $s^{(i)}$ merely contains a collection of lengths, which are summed up to give the next prefix. 
\end{proof}

It turns out that this same argument holds in general.

\begin{theorem}
Let $\{s^0\} = a$ for some $a \in A$ where $a \neq 1$, and let $\{s^{(i)}\}$ for $i \geq 1$ be the unique sequence that starts with $a$ and is such that its block lengths, read from left to right, reproduce the sequence $\{s^{(i-1)}\}$. Then the limit of this process, $s^\infty$, is the sequence $s(A,a)$.
\end{theorem}

Note that we must specify that $a \neq 1$, because that is the only case in which the length of $s^{(i+1)}$ is not strictly greater than the length of $s^{(i)}$. The proof of this theorem is omitted, as it is essentially the same proof as for the $A = \{1,2\}, a = 2$ case.

It turns out that we can be less restrictive about what the starting sequence is. Going back to the case of $\{1,2\}$, we can let $s^{(0)}$ be \emph{any} sequence of 1's and 2's that starts with a 2. For example, suppose $s^{(0)} = 2,1,1,1,2$. The rules for generating large $s^{(i)}$ give:
\begin{align*}
\{s^{(0)}\} &= \mathbf{2},1,1,1,2\\
\{s^{(1)}\} &= \mathbf{2,2,1},2,1,2,2\\
\{s^{(2)}\} &= \mathbf{2,2,1,1,2,1},1,2,1,1,2,2\\
\{s^{(3)}\} &= \mathbf{2,2,1,1,2,1,2,2,1,2},1,1,2,1,2,2,1,1\\
&\cdots
\end{align*}
Here, we highlight in bold the numbers that form a prefix of $s(\{1,2\},2)$. We note that $s^\infty = s(A,a)$ simply because $s^{(0)}$ contains a 2 at the beginning, whch then causes a 2,2 to appear at the start of $s^{(1)}$, which causes $2,2,1,1$ to appear at the start of $s^{(2)}$, and so on. The rest of each $s^{(i)}$ that is not a prefix of $s({1,2},2)$ we call the \emph{tail}.

We now consider two different forms of convergence. We know that $s^{(i)}$ in some sense approaches $s(A,a)$, as arbitrarily many elements at the beginning of the sequence will match up with the beginning of $s(A,a)$ as $i$ increases. However, if the tail of the sequence never entirely disappears, then there does not exist an $i$ where $s^{(i)}$ is exactly a prefix of $s(A,a)$. Thus, we can formalize two types of convergence: \emph{weak convergence} and \emph{strong convergence}. Weak convergence is satisfied if $s^\infty = s(A,a)$, and strong convergence is satisfied if $s^{(i)}$ is a prefix of $s(A,a)$ for all $i \geq k$ for some $k$.

@@

So we can ask, when can we do this? We would like to be able to simulate our original algorithm with this current one. We had established that our algorithm had a state corresponding to 
\begin{itemize}
\item current block length number
\item how many elements already written down in the block
\item which number we need to write.
\end{itemize}

In an algorithm with rules, we cannot remember anything about the previous rules. Therefore, each rule needs to encompas the block length numbers, the number of elements already written, and which number we need to write. The first condition is trivial to keep, since the rules themselves will say the block length numbers, in order to keep track of the numbers already written, each rule should be self-contained. That means that the rules should read off the block numbers. We would also need the length of the rules to be divisible by the size of the generating set. This is so that at the start of each rule, we always know which numbers to write.

Therefore, we can formulate a set of sufficient conditions to make rules to generate the self-referential sequence with generating set $G$. We want to have a set of rules $R$ where each rule $r \in R$ has length $|r| = l$, such that all $|G|^{l}$ strings are in the rule set and for each $r \in R$
\begin{itemize}
\item each rule reads off the block lengths in the same order of $G$
\item $\sum_{i \in r} i$ is divisible by $l$
\item $l$ is divisible by $|G|$
\end{itemize}
Given these conditions on every rule, we can guarantee that the self-referential sequence is generated. Note that in this case, the above example does not satisfy our rules until we say that $1, 3 \to 3, 1, 1, 1$.

Another exaple, we can have $G = \{ 2, 4 \}$, and we can have that the rules be
\begin{align*}
\text{Rule }A: 2,2 &\to 2, 2, 4, 4 \\
\text{Rule }B: 2,4 &\to 2, 2, 4,4,4,4\\
\text{Rule }C: 4,2 &\to 2,2,2,2,4,4\\
\text{Rule }D: 4,4 &\to 2, 2, 2, 2, 4, 4, 4, 4
\end{align*}

@@

We believe that if there is a way to show this, we will be able to show that the density is $\frac{1}{2}$, since the density will be bounded below by $\sum_{n=1}^\infty \dfrac{1}{3^n} = \frac{1}{2}$ and bounded above by the same number. However, generalizing the procedure to improve the bound seems dificult.

\end{document}
